---
title: "SetupProject"
date: "`r Sys.Date()`"
output:
  html_vignette:
    toc: true
    toc_depth: 2
    keep_md: true
vignette: >
  %\VignetteIndexEntry{SetupProject}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  results = 'asis'
)
```

# Introduction

The `start_new_reporting` function helps set up a structured Quarto document for statistical reporting. It generates a folder containing necessary files for report creation, including Quarto files (`.qmd`), custom CSS for styling, an HTML header, and a configuration YAML file.

# Function Definition

This function creates and writes a Quarto document (`.qmd`), along with custom CSS and HTML headers, to structure a statistical analysis report. It takes the following arguments:

- `name` The name of the quarto document to be generated.
- `folder_path` The folder where the project files should be created.
- `output_folder` The folder where the HTML report will be stored (a local copy is always kept, just in case).
- `structure` Character string indicating the organizational structure, either "USMR" or "EUCLID". Default is "USMR".
- `path_logo` Character string specifying the path to the logo image. The package contains the logo for EUCLID and USMR, so you can just put `NULL` there.
- `confidential` Logical value indicating whether the report should be marked as confidential. Default is FALSE.
- `report_type` Character string specifying the type of report. Default is "Data review report".
- `study_id` Character string representing the study identifier. Default is "CHUBXYYYY/NN".
- `study_name` Character string specifying the name of the study. Default is "The Study Name".
- `study_abbreviation` Character string providing the abbreviation of the study. Default is "TSN".
- `investigator` Character string representing the investigator's name. Default is "Investigator name".
- `methodologist` Character string specifying the methodologist's name. Default is "Jean Dupont".
- `biostatistician` Character string specifying the biostatistician's name. Default is "George Frais".
- `datamanager` Character string specifying the data manager's name. Default is "Peter Parker".
- `methodologist_mail`, `biostatistician_mail`, `datamanager_mail`: The emails of the different persons, you can also let it guess and put NULL there.
- `language` Character string indicating the language of the report, either "fr" (French) or "en" (English). Default is "fr".

Using all those arguments, the fonction generate several files that you can modify afterwards anyway.

# Generated Files

The function generates a folder containing several essential files:

- **`<name>.qmd`**: The main Quarto document for statistical analysis.
- **`<name>_custom.css`**: A custom CSS file to style the generated HTML report.
- **`<name>_custom.html`**: An HTML header file included before the document body.
- **`_quarto.yml`**: A Quarto configuration YAML file defining the report setup.
- **`datestamp_output_file.R`**: An R script used for post-processing the HTML report and saving it in the appropriate location.

# Example

```{r eval = FALSE}
start_new_reporting(folder_path = "reporting/",
                    output_folder = "UnEndroitSurleReseau/",
                    name = "report_analysis",
                    structure = "USMR",
                    confidential = FALSE,
                    report_type = "Data review report",
                    study_id = "CHUBXYYYY/NN",
                    study_name = "Chocolatines vs Pain au chocolat",
                    study_abbreviation = "CPC",
                    investigator = "Croissant AuBeurre",
                    methodologist = "Boulanger Bordelais",
                    biostatistician = "Patissier Toulousain",
                    datamanager = "Meunier Palois",
                    language = "fr")
```


# Conclusion

Using `start_new_reporting` ensures a structured and standardized approach to statistical reporting, making it easier to maintain consistency and readability in data analysis documents.

